NanoMaLy

### 2015-11-19 paper/pldi16-submission/interactive.tex
remove text from interactive.tex

\section{Explaining Type Errors With Traces}
\label{sec:explaining}
%
We have shown how to reliably find witnesses to type errors in \ocaml,
but this not fully address our original goal -- to \emph{explain} the
errors.
%
Having identified an input vector that triggers a bug, a common next
step is to step through the program with a \emph{debugger} to observe
how the program evolves.
%
The existing debuggers and interpreters for \ocaml assume a type-correct
program, so unfortunately we cannot use them off-the-shelf.
%
Instead we return to our semantics for \lang and extend the evaluation
rules to collect a trace that we can present to the user, demonstrating
precisely \emph{how} their program went wrong.

The trace takes the form of a \emph{reduction graph}, where the nodes
are terms and the edges represent either the single-step
$\hookrightarrow$ or a ``sub-term'' relation. For example, evaluating
the expression @1 + 2 + 3@ would produce the graph in
Figure~\ref{fig:simple-reduction}.
%
\begin{figure}[t]
  \centering
  \includegraphics[width=\linewidth]{simple.png}
\caption{The reduction graph for \texttt{1 + 2 + 3}.}
\label{fig:simple-reduction}
\end{figure}
%
We choose this graph representation instead of a simple, linear sequence
of expressions because it will allow us to express a variety of
traversals, such as ``step into'' and ``step over'' -- commonly found in
traditional debuggers.

\subsection{Interactive Semantics}
\label{sec:inter-semant}
%
\begin{figure*}[t]
\relDescription{Trace Syntax}
$$
\begin{array}{rrcl}
  & \tr & ::= & \bullet \spmid \singlestep{e}{e}; \tr \spmid \subterm{e}{e}; \tr
\end{array}
$$
\\
\relDescription{\subtermssym}
\begin{gather*}
\begin{array}{lcl}
\subterms{\eapp{e_1}{e_2}}   & \defeq & \subterm{\eapp{e_1}{e_2}}{e_1}; \subterm{\eapp{e_1}{e_2}}{e_2} \\
\subterms{\eplus{e_1}{e_2}}   & \defeq & \subterm{\eplus{e_1}{e_2}}{e_1}; \subterm{\eplus{e_1}{e_2}}{e_2} \\
\subterms{\eif{e_1}{e_2}{e_3}}   & \defeq & \subterm{\eif{e_1}{e_2}{e_3}}{e_1}; \\
                                &        & \subterm{\eif{e_1}{e_2}{e_3}}{e_2}; \\
                                &        & \subterm{\eif{e_1}{e_2}{e_3}}{e_3} \\
\subterms{\elet{x}{e_1}{e_2}}   & \defeq & \subterm{\elet{x}{e_1}{e_2}}{e_1}; \\
                                &        & \subterm{\elet{x}{e_1}{e_2}}{e_2} \\
\subterms{\efun{x}{e}}       & \defeq & \subterm{\efun{x}{e}}{e} \\
\subterms{e}                 & \defeq & \bullet
\end{array}
\end{gather*}
\judgementHead{Evaluation}{\stepg{e}{\su}{\tr}{e}{\su}{\tr}}
\begin{gather*}
\inference[\recontext]
  {\stepg{e}{\su}{\tr}{e_1}{\su_1}{\tr_1}}
  {\stepg{C[e]}{\su}{\tr}{C[e_1]}{\su_1}{\singlestep{C[e]}{C[e_1]}; \subterms{C[e_1]}; \tr_1}}
\\ \\
\inference[\reappgood]
  {\pair{\efun{x}{e}}{\su_2} = \force{v_1}{\tfun{\thole{}}{\thole{}}}}
  {\stepg{\eapp{v_1}{v_2}}{\su_1}{\tr}
         {e\sub{x}{v_2}}{\su_1;\su_2}{\singlestep{\eapp{v_1}{v_2}}{e\sub{x}{v_2}}; \subterms{e\sub{x}{v_2}}; \tr}}
\\ \\
\inference[\reappbad]
  {\pair{\stuck}{\su_2} = \force{v_1}{\tfun{\thole{}}{\thole{}}}}
  {\stepg{\eapp{v_1}{v_2}}{\su_1}{\tr}{\stuck}{\su_1;\su_2}{\singlestep{\eapp{v_1}{v_2}}{\stuck}; \tr}}
\end{gather*}
\caption{A selection of the operational semantics from
  Figure~\ref{fig:operational}, extended to collect a full reduction
  graph.}
\label{fig:interactive}
\end{figure*}
%
The changes to the operational semantics of \S\ref{sec:semantics} are
mechanical, so we will not reproduce them in full; instead we describe
the procedure for extending a transition rule and provide a selection of
examples in Figure~\ref{fig:interactive}.

@@

>>>>

\section{Explaining Type Errors With Traces}
\label{sec:explaining}
%
% We have shown how to reliably find witnesses to type errors in \ocaml,
% but this not fully address our original goal -- to \emph{explain} the
% errors.
% %
% Having identified an input vector that triggers a bug, a common next
% step is to step through the program with a \emph{debugger} to observe
% how the program evolves.
% %
% The existing debuggers and interpreters for \ocaml assume a type-correct
% program, so unfortunately we cannot use them off-the-shelf.
% %
% Instead we return to our semantics for \lang and extend the evaluation
% rules to collect a trace that we can present to the user, demonstrating
% precisely \emph{how} their program went wrong.

% The trace takes the form of a \emph{reduction graph}, where the nodes
% are terms and the edges represent either the single-step
% $\hookrightarrow$ or a ``sub-term'' relation. For example, evaluating
% the expression @1 + 2 + 3@ would produce the graph in
% Figure~\ref{fig:simple-reduction}.
% %
% \begin{figure}[t]
%   \centering
%   \includegraphics[width=\linewidth]{simple.png}
% \caption{The reduction graph for \texttt{1 + 2 + 3}.}
% \label{fig:simple-reduction}
% \end{figure}
% %
% We choose this graph representation instead of a simple, linear sequence
% of expressions because it will allow us to express a variety of
% traversals, such as ``step into'' and ``step over'' -- commonly found in
% traditional debuggers.

In this section we formalize the collection and traversal of the reduction graph.
%
First, we extend our semantics from \S~\ref{sec:semantics} to collect
the edges in the graph (\S~\ref{sec:inter-semant}).
%
Then, we express a set of common debugging steps as graph traversals
(\S~\ref{sec:traversing-graph}).

\subsection{Tracing Semantics}
\label{sec:inter-semant}
%
\begin{figure*}[t]
\relDescription{Trace Syntax}
$$
\begin{array}{rrcl}
  & \tr & ::= & \bullet \spmid \singlestep{e}{e}; \tr \spmid \subterm{e}{e}; \tr
\end{array}
$$
\\
\relDescription{\subtermssym}
\begin{gather*}
\begin{array}{lcl}
\subtermssym                 & \dcolon & e \to \tr \\
\subterms{\eapp{e_1}{e_2}}   & \defeq & \subterm{\eapp{e_1}{e_2}}{e_1}; \subterm{\eapp{e_1}{e_2}}{e_2} \\
\subterms{\eplus{e_1}{e_2}}   & \defeq & \subterm{\eplus{e_1}{e_2}}{e_1}; \subterm{\eplus{e_1}{e_2}}{e_2} \\
\subterms{\eif{e_1}{e_2}{e_3}}   & \defeq & \subterm{\eif{e_1}{e_2}{e_3}}{e_1}; \\
                                &        & \subterm{\eif{e_1}{e_2}{e_3}}{e_2}; \\
                                &        & \subterm{\eif{e_1}{e_2}{e_3}}{e_3} \\
\subterms{\elet{x}{e_1}{e_2}}   & \defeq & \subterm{\elet{x}{e_1}{e_2}}{e_1}; \\
                                &        & \subterm{\elet{x}{e_1}{e_2}}{e_2} \\
\subterms{\efun{x}{e}}       & \defeq & \subterm{\efun{x}{e}}{e} \\
\subterms{e}                 & \defeq & \bullet
\end{array}
\end{gather*}
\judgementHead{Evaluation}{\stepg{e}{\su}{\tr}{e}{\su}{\tr}}
\begin{gather*}
\inference[\recontext]
  {\stepg{e}{\su}{\tr}{e_1}{\su_1}{\tr_1}}
  {\stepg{C[e]}{\su}{\tr}{C[e_1]}{\su_1}{\singlestep{C[e]}{C[e_1]}; \subterms{C[e_1]}; \tr_1}}
\\ \\
\inference[\reappgood]
  {\pair{\efun{x}{e}}{\su_2} = \force{v_1}{\tfun{\thole{}}{\thole{}}}{\su_1}}
  {\stepg{\eapp{v_1}{v_2}}{\su_1}{\tr}
         {e\sub{x}{v_2}}{\su_1;\su_2}{\singlestep{\eapp{v_1}{v_2}}{e\sub{x}{v_2}}; \subterms{e\sub{x}{v_2}}; \tr}}
\\ \\
\inference[\reappbad]
  {\pair{\stuck}{\su_2} = \force{v_1}{\tfun{\thole{}}{\thole{}}}{\su_1}}
  {\stepg{\eapp{v_1}{v_2}}{\su_1}{\tr}{\stuck}{\su_1;\su_2}{\singlestep{\eapp{v_1}{v_2}}{\stuck}; \tr}}
\end{gather*}
\caption{A selection of the operational semantics from
  Figure~\ref{fig:operational}, extended to collect a full reduction
  graph.}
\label{fig:interactive}
\end{figure*}
%
The changes to the operational semantics of \S~\ref{sec:semantics} are
mechanical, so we will not reproduce them in full; instead we describe
the procedure for extending a transition rule and provide a selection of
examples in Figure~\ref{fig:interactive}.

@@

After evaluation the reduction graph can be constructed directly from
the trace $\tr$ as follows:
\[
G(\tr) = \pair{\{e \spmid e \in \tr\}}{\tr}
\]
